### 2022-01-08 cleave.tex
fixing

\todo[inline]{
    For the approach part, mention please also that it is open-source and where it can be found. Also, you keep the discussion around the implemented plant and the controller pretty short (and it is only showing up in the numerical part). Maybe it is better to discuss the details of the implemented plant and the controller already in Section 2 and mention that the intention is to build a liberary of plants and controllers in the future.
}

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.

@@

>>>>

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.

@@

\subsection{Implemented Control Loops}

\begin{figure}
    \centering
    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
    \caption{
        The two-dimensional inverted pendulum system.
        The cart moves on the X-axis, and the pendulum on top of it swings freely.
        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
    }\label{fig:invpend}
\end{figure}

For the purpose of this work, we have used \ac{CLEAVE} to implement an emulated \emph{inverted pendulum} control loop; see \cref{fig:invpend}.
We have chosen this system as an initial benchmark for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
However, we eventually plan to build an open library of emulated control loops to be able to provide more thorough and varied benchmarks.

The inverted pendulum Plant is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
For the Controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.

### 2022-01-08 experiments.tex
fixing

\todo[inline]{
    In the numerical part, you can gain a lot of space by reformatting the plots into environments with two or three plots next to each other. In addition, you need to comment more on Fig 4 (the key result for the single-loop baseline). The discussion of Fig 4 should take more space than the discussion of Fig 5.
}

@@

The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.

\begin{figure}
    \centering
    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
    \caption{
        The two-dimensional inverted pendulum system.
        The cart moves on the X-axis, and the pendulum on top of it swings freely.
        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
    }\label{fig:invpend}
\end{figure}

@@

\begin{figure*}[h]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_toppled}
        \caption{}%
        \label{fig:single:topple}
    \end{subfigure}%
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_rms}
        \caption{}\label{fig:single:rms}
    \end{subfigure}%
    \caption[caption]{
        Stability metrics for the single-loop scenarios.
        \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
        }%
    \label{fig:single:stability}
    \todo[inline]{More focus on this.}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{single_loop_rtts}
    \caption{
        Mean latency due to network and processing for both single-loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
    }\label{fig:single:rtt}
    \todo[inline]{less focus on this. mention packet drops are all below 0.2 percent}
\end{figure}

@@

We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
\Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the penduli angles.

We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.

Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.

From these results we can already infer interesting consequences for \ac{NCS} deployments on the edge.
For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.

@@

\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
These results are interesting in their counter-intuitiveness.
Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.

\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.

>>>>

@@

@@

\begin{figure*}[h]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_toppled}
        \caption{}%
        \label{fig:single:topple}
    \end{subfigure}%
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_rms}
        \caption{}\label{fig:single:rms}
    \end{subfigure}%
    \caption[caption]{
        Stability metrics for the single-loop scenarios.
        \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
        }%
    \label{fig:single:stability}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{single_loop_rtts}
    \caption{
        Mean latency due to network and processing for both single-loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
    }\label{fig:single:rtt}
\end{figure}

@@

We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable collection and posterior processing of relevant metrics.  
\Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the penduli angles.
We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.

As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
As sampling rates increased, the system was able to reach stability at higher \ac{RTT} latencies.
These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.

For completeness, we also showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.

@@

\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
These results are interesting in their counter-intuitiveness.
Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
